An Efficient Feedback Calibration Algorithm for Direct Imaging Radio Telescopes

### 2016-06-22 cal_demo/ms.tex
point to mnras.cls in parent directory
Added "aperture" before all cases of "illumination pattern"

% Basic setup. Most papers should leave these options alone.
\documentclass[a4paper,fleqn,usenatbib]{mnras}

@@

The next step is to grid the calibrated electric field measurements to a finite set of regularly spaced 
positions on the 2D plane of the ground, $\mathbf{r}_j$.
The extent of the grid is chosen to encompass the footprint of the antenna array.
MOFF uses the antenna illumination patterns as the gridding 
kernel according to the software holography/A-transpose technique \citep{mor09,bha08}. 
Mathematically we can express the gridding operation as
\begin{equation}\label{eq:gridding}
E'_j=\sum_a W_a(\mathbf{r}_j-\mathbf{r}_a) E'_a(t)
\end{equation}
Formally each pixel is calculated by summing over all antennas and their illumination patterns.
However the illumination patterns are typically compact and non-overlapping, so most pixels
will only contain contribution from at most one antenna, allowing equation~\ref{eq:gridding}
to be calculated through a sparse matrix multiplication.
This gridding 
achieves an optimal map \citep{teg97b} and places the data on a regular grid. Unlike 
traditional correlators which grid visibilities, the MOFF grids the electric field measurements directly. This 
operation must be performed for every time sample, before any averaging of the data.

@@

The pipeline up to this point can be summarized 
through the following equation.
\begin{align}
\mathcal{E}'(\s_k,t) & = \underbrace{\frac{1}{\Nant} \sum_j e^{2\pi i \mathbf{r}_j \cdot \s_k}}_{\mathrm{2D\;FFT}} 
\underbrace{\sum_a W_a(\mathbf{r}_j - \ra)}_{\mathrm{Gridding}} \\
& \qquad \times \underbrace{ \frac{1}{g'_a} \left(g_a E_a^I(t)+n_a(t)\right)}_{\mathrm{Calibration}} \nonumber
\end{align}
The sum over antennas uses the antenna illumination pattern to grid the calibrated electric field measurements onto 
regular gridpoints, $\mathbf{r}_j$. The sum over $j$ denotes the 2D FFT to sky coordinates, resulting in an 
estimate for the instantaneous electric field image. We can simplify this expression by 
exchanging the sums and use the DFT to approximate the Fourier relationship between
the antenna illumination pattern and the far-field radiation pattern.
\begin{align}\label{eq:epix}
\mathcal{E}'(\s_k,t) & = \frac{1}{\Nant} \sum_a \frac{1}{g'_a}\left(g_a E^I_a(t)+n_a(t)\right) e^{2\pi i \s_k \cdot \ra} \nonumber \\
  & \qquad\times \sum_j W_a(\mathbf{r}_j-\ra)e^{2\pi i \s_k \cdot (\mathbf{r}_j-\ra)} \nonumber\\
& \approx \frac{1}{\Nant} \sum_a \frac{1}{g'_a}\left(g_aE^I_a(t)+n_a(t)\right) e^{2\pi i \s_k \cdot \ra}\mathcal{W}_a(\s_k)
\end{align}
In this form we see that the effect of gridding with the antenna illumination pattern is to attenuate 
the image by a factor of the far-field radiation response, $\mathcal{W}_a(\s)$.

@@

The physical parameters of our simulation can be arbitrarily scaled to other cases. However, 
we provide typical units for an MWA observation as an example. Our simulated antenna 
array consists of the inner 51 antennas of the MWA layout \citep{bea12}, within a bounding 
box of 150~m. The antenna illumination pattern used is a 4.4~m square tophat; a very rough 
approximation to the square shaped tiles of the MWA. Because our algorithm treats frequency 
channels independently, we simulate only one channel. For context we treat this as a channel
at 150~MHz with 40 kHz bandwidth, and adopt a sampling period of 25~$\mu$s. The simulated signal 
consists of 10 random point sources with flux densitities 0.5~Jy~$\lesssim S \lesssim$~1~Jy 
within the main lobe of the primary beam.

>>>>

% Basic setup. Most papers should leave these options alone.
\documentclass[a4paper,fleqn,usenatbib]{../mnras}

@@

The next step is to grid the calibrated electric field measurements to a finite set of regularly spaced 
positions on the 2D plane of the ground, $\mathbf{r}_j$.
The extent of the grid is chosen to encompass the footprint of the antenna array.
MOFF uses the antenna aperture illumination patterns as the gridding 
kernel according to the software holography/A-transpose technique \citep{mor09,bha08}. 
Mathematically we can express the gridding operation as
\begin{equation}\label{eq:gridding}
E'_j=\sum_a W_a(\mathbf{r}_j-\mathbf{r}_a) E'_a(t)
\end{equation}
Formally each pixel is calculated by summing over all antennas and their aperture illumination patterns.
However the aperture illumination patterns are typically compact and non-overlapping, so most pixels
will only contain contribution from at most one antenna, allowing equation~\ref{eq:gridding}
to be calculated through a sparse matrix multiplication.
This gridding 
achieves an optimal map \citep{teg97b} and places the data on a regular grid. Unlike 
traditional correlators which grid visibilities, the MOFF grids the electric field measurements directly. This 
operation must be performed for every time sample, before any averaging of the data.

@@

The pipeline up to this point can be summarized 
through the following equation.
\begin{align}
\mathcal{E}'(\s_k,t) & = \underbrace{\frac{1}{\Nant} \sum_j e^{2\pi i \mathbf{r}_j \cdot \s_k}}_{\mathrm{2D\;FFT}} 
\underbrace{\sum_a W_a(\mathbf{r}_j - \ra)}_{\mathrm{Gridding}} \\
& \qquad \times \underbrace{ \frac{1}{g'_a} \left(g_a E_a^I(t)+n_a(t)\right)}_{\mathrm{Calibration}} \nonumber
\end{align}
The sum over antennas uses the antenna aperture illumination pattern to grid the calibrated electric field measurements onto 
regular gridpoints, $\mathbf{r}_j$. The sum over $j$ denotes the 2D FFT to sky coordinates, resulting in an 
estimate for the instantaneous electric field image. We can simplify this expression by 
exchanging the sums and use the DFT to approximate the Fourier relationship between
the antenna aperture illumination pattern and the far-field radiation pattern.
\begin{align}\label{eq:epix}
\mathcal{E}'(\s_k,t) & = \frac{1}{\Nant} \sum_a \frac{1}{g'_a}\left(g_a E^I_a(t)+n_a(t)\right) e^{2\pi i \s_k \cdot \ra} \nonumber \\
  & \qquad\times \sum_j W_a(\mathbf{r}_j-\ra)e^{2\pi i \s_k \cdot (\mathbf{r}_j-\ra)} \nonumber\\
& \approx \frac{1}{\Nant} \sum_a \frac{1}{g'_a}\left(g_aE^I_a(t)+n_a(t)\right) e^{2\pi i \s_k \cdot \ra}\mathcal{W}_a(\s_k)
\end{align}
In this form we see that the effect of gridding with the antenna aperture illumination pattern is to attenuate 
the image by a factor of the far-field radiation response, $\mathcal{W}_a(\s)$.

@@

The physical parameters of our simulation can be arbitrarily scaled to other cases. However, 
we provide typical units for an MWA observation as an example. Our simulated antenna 
array consists of the inner 51 antennas of the MWA layout \citep{bea12}, within a bounding 
box of 150~m. The antenna aperture illumination pattern used is a 4.4~m square tophat; a very rough 
approximation to the square shaped tiles of the MWA. Because our algorithm treats frequency 
channels independently, we simulate only one channel. For context we treat this as a channel
at 150~MHz with 40 kHz bandwidth, and adopt a sampling period of 25~$\mu$s. The simulated signal 
consists of 10 random point sources with flux densitities 0.5~Jy~$\lesssim S \lesssim$~1~Jy 
within the main lobe of the primary beam.
